---
title: "Deision trees"
author: "Camille"
date: "17/06/2017"
output: html_document
---

Loading the required packages
```{r}
load_packages  <- function(p){
  
  for(i in seq_along(p)) {
    if(!require(p[i], character.only=TRUE)) {
      install.packages(p[i])}
      library(p[i], character.only=TRUE)
  }

}

list.of.packages <- c("rpart","lubridate","outliers","rpart.plot", "xgboost",
                      "caret","caretEnsemble", "randomForest","e1071","pROC", "tidyr", "klaR", 
                      "car","devtools","yamldebugger","mlbench","Hmisc", "ggvis","relaimpo")

sapply(list.of.packages,load_packages)
```

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading train and test sets
```{r}
train <- readRDS("train.b1.v1")
test <- readRDS("test.b1.v1")
```


#### Full decision tree ####

```{r}
dt.full <- rpart(Sales ~ ., data = train) ## Using default parameter values
pred.dt <- predict(dt.full,newdata=test)

printcp(dt) ## Find the cp value with the least relative error
```


### Training the tree using caret to find the optimal value of cp

```{r}
dt.ctrl <- trainControl(method = "cv",number=2,repeats = 0,
                        savePredictions =T) # problem with missing values

dt.caret <- train(Sales~.,data=train,method="rpart",
                  trControl=dt.ctrl)

summary(dt.caret)
dt.caret$pred
pred.dt <- predict(dt.caret,newdata=test)
```


# Optimal value of cp=??? need to determine the cp value

### Manual training

Caret only allows us to modify the cp value, manual training was done to find the optimal minsplit

```{r}
dt.param <- expand.grid(minsplit=c(1:50),cp=seq(0.00001,6.0749e-04,0.00001))
k <- 5
results.dt <- matrix(data=NA, nrow =k ,ncol=nrow(dt.param))
folds <- cut(1:nrow(train),breaks = k,labels = F)

for(j in 1:nrow(dt.param)){
  for(i in 1:k){
    
    cv.idx <- which(folds==i)
    cv.val <- train[cv.idx,]
    cv.train <- train[-cv.idx,] 
    dt.train <- rpart(Sales ~ ., data = cv.train,
                      minsplit=dt.param[j,1],cp=dt.param[j,2])  
    
    pred.dt  <- predict(dt.caret,newdata=cv.val)
    
  }}


```

## Final model - Optimal parameters 

```{r}
dt <- rpart(Sales ~ ., data = train,cp=0.00046 ,xval=10)

pred.dt <- predict(dt,newdata=test)
pred.dt$confusion_matrix
pred.dt$finalscore

printcp(dt)
summary(dt)
rpart.plot(dt)
```


## Trying dt with loss function

```{r}
dt.loss <- rpart(Sales ~ .,data=train,parms=list(split="information", loss=matrix(c(0,10,3,0), byrow=TRUE, nrow=2)),minsplit=47,cp=0.00046)

pred.dt.loss <- predict(dt.loss,newdata=test)
pred.dt.loss$finalscore

```

