---
title: "Data preparation"
author: "Group-9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data preparation ## 

The following code loads the respective datasets and reformats the variables into the appropriate datatype


Loading required packages 
```{r}
if (!require("rstudioapi")) install.packages("rstudioapi"); library("rstudioapi")
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

source("Helperfunction.R")

list.of.packages <- c("rpart","lubridate","outliers","rpart.plot", "xgboost",
                      "caret","randomForest","e1071","pROC", "tidyr", "klaR", 
                      "caretEnsemble","car","mlbench","Hmisc", "ggvis", "mlr")

sapply(list.of.packages,load_packages)
```

Loading the datasets 
```{r}
# Specify file path and parts of the file name to match

data.list<-load.multiple.files(path=getwd(),pattern = "assignment_BADS_WS1617_*csv*")

class1 <-data.list$assignment_BADS_WS1617_class.csv
known1 <-data.list$assignment_BADS_WS1617_known.csv

```

Replace apparent typos - replace year "4746" with "2013"(mode) 
```{r}
# Class dataset
class1$deliverydate_estimated <- gsub("4746", "2013", class1$deliverydate_estimated)

# Known dataset
known1$deliverydate_estimated <- gsub("4746", "2013", known1$deliverydate_estimated)
known1$deliverydate_actual <- gsub("4746", "2013", known1$deliverydate_actual)
```

Reformat variables to the correct data type
```{r}
# Date variables
date_var <- c("order_date","account_creation_date","deliverydate_actual","deliverydate_estimated")
class1[,date_var] <- lapply(class1[,date_var], ymd)

known1[,date_var] <- lapply(known1[,date_var], ymd)

# Factor variables
varlist <- c("ID","title","newsletter","model","delivery","coupon","giftwrapping","referrer","cost_shipping")
class1[,varlist] <- lapply(class1[,varlist],factor)
sapply(class1,class)

known1[,varlist] <-lapply(known1[,varlist],factor)
sapply(known1,class)

known1$return_customer <- factor(known1$return_customer) 
```

Delete empty/unnecessary columns (points_redeemed, ID, hardware_count) 
```{r}
# Class dataset
class1$points_redeemed<-NULL
class1$hardware_count <- NULL
class1$ID <- NULL

# Known dataset
known1$points_redeemed<-NULL
known1$hardware_count <- NULL
known1$ID <- NULL
```

This script cleans the dataset by removing outliers and NA values 

## Treatment of missing values ##

Create dummy variables for variables with missing values
```{r}
# missing variables
class1[,c("deliverydate_estimated_miss",
          "weight_miss")] <-lapply(class1[,c("deliverydate_estimated","weight")],is.na)

known1[,c("deliverydate_estimated_miss",
          "weight_miss")] <-lapply(known1[,c("deliverydate_estimated","weight")],is.na)
```

## Check variance of predictors ##
```{r}
# Check for near zero variance predictors using function from caret package
nzv <- nearZeroVar(known1, saveMetrics= TRUE)
print(nzv)

# Variable "deliverydate_estimated_miss" appears to be a near zero variance predictor and should be dropped
class1$deliverydate_estimated_miss <- NULL
known1$deliverydate_estimated_miss <- NULL
```

## Replace missing values in existing variables ##

Replace NAs in form_of_address
```{r}
# Class dataset
class1$form_of_address <- factor(class1$form_of_address, levels=c(levels(class1$form_of_address), "missing"))
form_na <- is.na(class1$form_of_address)
class1[form_na,"form_of_address"]<- "missing"

# Known dataset
known1$form_of_address <- factor(known1$form_of_address, levels=c(levels(known1$form_of_address), "missing"))
form_na <- is.na(known1$form_of_address)
known1[form_na,"form_of_address"]<- "missing"
```

Replace NA in deliverydate_actual with deliverydate_estimated
```{r}
# Class dataset
missing_delivery_c <- class1[is.na(class1$deliverydate_actual),]
missing_delivery_c$deliverydate_actual <- missing_delivery_c$deliverydate_estimated
class1[is.na(class1$deliverydate_actual), 
       "deliverydate_actual"] <- missing_delivery_c$deliverydate_actual

# Known dataset
missing_delivery_k <- known1[is.na(known1$deliverydate_actual),]
missing_delivery_k$deliverydate_actual <- missing_delivery_k$deliverydate_estimated
known1[is.na(known1$deliverydate_actual), 
       "deliverydate_actual"] <- missing_delivery_k$deliverydate_actual
```

Replacing the missing values in postcode_delivery 
```{r}
# Class dataset
class1$postcode_delivery <- as.integer(class1$postcode_delivery)
class1$postcode_invoice <- as.integer(class1$postcode_invoice)

my.na_class<-is.na(class1$postcode_delivery)
class1$postcode_delivery[my.na_class]<-class1$postcode_invoice[my.na_class] 

# Known dataset
known1$postcode_delivery <- as.integer(known1$postcode_delivery)
known1$postcode_invoice <- as.integer(known1$postcode_invoice)

my.na_known<-(known1$postcode_delivery=="1")
known1$postcode_delivery[my.na_known]<-known1$postcode_invoice[my.na_known]
```

Replace NAs in advertising_code
```{r}
# Class dataset
levels(class1$advertising_code) <- sub("^$", "default", levels(class1$advertising_code))

# Known dataset
levels(known1$advertising_code)<-sub("^$", "default", levels(known1$advertising_code))
```

Impute weight variable with the mode value (=0) and convert units into kg
```{r}
options(max.print=52000) #This may be needed to run the following

# Known dataset
countweightk <- table(known1$weight) # Checking the distribution of the weight
names(countweightk)[countweightk == max(countweightk)] # Getting the mode
known1$weight[is.na(known1$weight)] <- 0 # Imputing the mode in place of NA's
known1$weight <- known1$weight/1000

# Class dataset
countweightc <- table(class1$weight) # Checking the distribution of the weight
names(countweightc)[countweightk == max(countweightk)] # Getting the mode
class1$weight[is.na(class1$weight)] <- 0 # Imputing the mode in place of NA's
class1$weight <- class1$weight/1000
```

## Check and remove outliers ##

We used the following code to check for and remove outliers. However, it did not have an
effect on the tree based models that we ultimately used.

Boxplots to check the distribution of variables
```{r}
#sapply(class1[, 24:37], boxplot)
```

Find the quartile values and the inter-quantile-range IQR for the examined variables
```{r}
#lower.quartile_i_k <- as.numeric(summary(known1$item_count)[2])
#upper.quartile_i_k <- as.numeric(summary(known1$item_count)[5])
#IQR_i_k <- upper.quartile_i_k - lower.quartile_i_k
```

Calculate the upper bound values
```{r}
#upper.bound_i_k <- upper.quartile_i_k + 1.5*IQR_i_k
#message("Upper bound on item_count is ", upper.bound_i_k )
```

Use logical indexing to identify outliers and replace with the upper bound
```{r}
# item_count
#known1$item_count[ known1$item_count > upper.bound_i_k ] <- upper.bound_i_k

# imported_count
#imported_out <-rm.outlier(known1$imported_count, fill = FALSE, 
                          #median = FALSE, opposite = FALSE)

# paperback_count
#paperback_o<-rm.outlier(known1$paperback_count, fill = FALSE, 
                        #median = FALSE, opposite = FALSE)

# schoolbook_count
#schoolbook_o<-rm.outlier(known1$schoolbook_count, fill = FALSE, 
                         #median = FALSE, opposite = FALSE)

# audiobook_count
#audiobook_o<- rm.outlier(known1$audiobook_count, fill = FALSE, 
                         #median = FALSE, opposite = FALSE)

# other_count
#other_o<-rm.outlier(known1$other_count, fill = FALSE, 
                    #median = FALSE, opposite = FALSE)
```

## Feature Engineering ##
#(create difference variables, separate date variables, binning, WOE)

Create delivery_date_diff variable
```{r}
class1$deliverydate_diff <- class1$deliverydate_actual - class1$deliverydate_estimated
known1$deliverydate_diff <- known1$deliverydate_actual - known1$deliverydate_estimated
```

Create dummy variable for whether customer has an account or not
```{r}
class1$account <- !is.na(class1$account_creation_date)
known1$account <- !is.na(known1$account_creation_date)
```

Create variable for length of account 
```{r}
class1$account_length <- class1$order_date - class1$account_creation_date
known1$account_length <- known1$order_date - known1$account_creation_date

# Impute NA values as "-1"
class1$account_length[is.na(class1$account_length)] <- -1
known1$account_length[is.na(known1$account_length)] <- -1
```

Convert difftime variables to integers
```{r}
known1$deliverydate_diff <- as.integer(known1$deliverydate_diff)
known1$account_length <- as.integer(known1$account_length)

class1$deliverydate_diff <- as.integer(class1$deliverydate_diff)
class1$account_length <- as.integer(class1$account_length)
```


Separate order_date, deliverydate_actual and deliverydate_estimated into year, month, day columns
```{r}
# Class dataset
class1 <- separate(class1, order_date, into = c("order_year", "order_month", 
                                                "order_day"), sep="-")

class1 <- separate(class1, deliverydate_actual, into = c("deliveryactual_year", 
                                                         "deliveryactual_month",
                                                         "deliveryactual_day"), sep="-")

class1 <- separate(class1, deliverydate_estimated, into = c("deliveryestimated_year",
                                                            "deliveryestimated_month",
                                                            "deliveryestimated_day"),
                   sep="-")

# Known dataset
known1 <- separate(known1, order_date, into = c("order_year", "order_month", 
                                                "order_day"), sep="-")

known1 <- separate(known1, deliverydate_actual, into = c("deliveryactual_year",
                                                         "deliveryactual_month",
                                                         "deliveryactual_day"), sep="-")

known1 <- separate(known1, deliverydate_estimated, into = c("deliveryestimated_year",
                                                            "deliveryestimated_month",
                                                            "deliveryestimated_day"), 
                   sep="-")

# Convert additional variables above from character back into numeric
date_num <- c("order_year","order_month","order_day",
              "deliveryactual_year","deliveryactual_month","deliveryactual_day", 
              "deliveryestimated_year","deliveryestimated_month","deliveryestimated_day")

known1[,date_num] <- lapply(known1[,date_num], as.integer)
class1[,date_num] <- lapply(class1[,date_num], as.integer)
```

Delete account_creation_date variable
```{r}
class1$account_creation_date <- NULL
known1$account_creation_date <- NULL
```

Bin postcode variables
```{r}
# Known dataset 
binningk<-function(num.vector){
  output<-cut(num.vector, breaks=c(0,10,20,30,40,50,60,70,80,90,100,104), include.lowest = TRUE, labels=paste0("level",1:11) )
  num.vector<-return(output)   
}

# Class dataset
binningc<-function(num.vector){
  output<-cut(num.vector, breaks=c(0,10,20,30,40,50,60,70,80,90,99), include.lowest = TRUE, labels=paste0("level",1:10) )
  num.vector<-return(output)   
}

# Class dataset
class1$postcode_invoice<-binningc(class1$postcode_invoice)
class1$postcode_delivery<-binningc(class1$postcode_delivery)

# Known dataset
known1$postcode_invoice<-binningk(known1$postcode_invoice)
known1$postcode_delivery<-binningk(known1$postcode_delivery)
```

Create empty factor variable in Class dataset
```{r}
#create class1$return_customer as a factor
class1$return_customer<- factor(0)
```

WOE for advertising_code
```{r}
# Known dataset
woe.object <- woe(return_customer ~ advertising_code, data =known1, zeroadj = 0.5)
known1[, c("woe.advertising_code")] <- woe.object$xnew
known1[, c("advertising_code")] <- NULL

# Class dataset
class1$advertising_code<-recode(class1$advertising_code,"'default'=-0.004913810;'AA'=0;'AB'= 0.210694778; 'AC'=-0.766731954; 'AD'=-0.766731954; 'AE'=0.270511389; 'AF'= 0.486031015; 'AG'=0.331880335; 'AH'=0.341096990; 'AI'=2.066481390; 'AJ'=-0.073584773; 'AK'=-0.160596150; 'AL'=-0.073584773; 'AM'=-1.459879134; 'AO'=0.149558778; 'AP'= 0.202202512; 'AQ'=0.486031015; 'AR'=0.965838644; 'AS'=-0.766731954; 'AT'= 0.006457935; 'AU'=-0.073584773; 'AV'=-0.543588402; 'AW'= 0.244868958; 'AX'= -0.890399617; 'AY' = -0.766731954; 'AZ' = 1.052426490; 'BA' = 0.486031015; 'BB' = 0.619562407; 'BC'=1.584643303; 'BD' = 0.057378025; 'BE' = -0.073584773; 'BF' = 1.484559845; 'BG' = -0.766731954; 'BH'= -0.073584773; 'BI'= 1.835957732; 'BK'= -0.361266846; 'BL'= 0.842705959; 'BM' = -0.543588402; 'BN'=-2.153026315; 'BO'=-0.218766783; 'BQ'= -0.074188820; 'BR'= -0.048478852; 'BS' =  0.619562407; 'BT'= -0.244856494; 'BU' = -0.766731954; 'BV'= -0.073584773; 'BW'=-0.766731954; 'BX'=-2.376169866; 'BY'=  1.025027516; 'BZ'= -0.147692745; 'CA'=  0.192891903",  as.factor.result=FALSE)

class1[, c("woe.advertising_code")] <- class1$advertising_code
class1[, c("advertising_code")] <- NULL
```

Save current dataset 
```{r}
saveRDS(class1, "class1_clean4.RDS")
saveRDS(known1, "known1_clean4.RDS")
```


Creating train and test datasets
```{r}
set.seed(123)

idx.train <- createDataPartition(y = known1$return_customer, p = 0.8, list = FALSE)

# Assign the train and test set by indices
train <- known1[idx.train, ]
test <-  known1[-idx.train, ]
```

Change factor levels to "yes" and "no"

```{r}
train$return_customer <- factor(train$return_customer, levels = c(0,1), labels = c("no", "yes"))
test$return_customer <- factor(test$return_customer, levels = c(0,1), labels = c("no", "yes"))
```

Saving test and train sets

```{r}
saveRDS(train, "train.RDS")
saveRDS(test, "test.RDS")
```

