---
title: "Random Forest"
author: "Group 9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### MODEL TRAINING USING CARET ####

### Loading the dataset ####


```{r}
train <- readRDS("train.RDS")
test <- readRDS("test.RDS")
```


Setting model control parameters 

```{r}
k <- 5
set.seed(123)

model.control <- trainControl(
  method = "cv", 
  number = k, 
  repeats=3,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

rf.parms <- expand.grid(mtry = 1:10)
```

Model training using caret
```{r}
rf.caret <- train(return_customer~., data = train,  
                  method = "rf", ntree =500, tuneGrid = rf.parms, 
                  metric = "ROC", trControl = model.control,verboseIter=T,do.trace=100)
```


Performance metrics and results 

```{r}
rf.caret$results
rf.caret$finalModel

plot(rf.caret)
```


Predict the outcomes of the test set with the custom results function

```{r}
res.rf.caret <- results(rf.caret)
```

Selecting important variables

```{r}
imp <- varImp(rf.caret)$importance
save(imp,file = "impvarcaret.RDS")

imp$variable <- rownames(imp)
bestvars <- as.vector(row.names(imp[imp$Overall>=mean(imp$Overall),]))

bestvars <- gsub("email_domaingmx.de","email_domain",bestvars)
bestvars <- gsub("form_of_addressMr","form_of_address",bestvars)
bestvars <- gsub("newsletter1","newsletter",bestvars)
bestvars <- bestvars[-c(5,6)]

frml_rfvars <- as.formula(paste("return_customer ~" ,paste(bestvars,collapse ="+"),sep=""))
```



### FINAL RANDOM FOREST ####

## Run random forest again using the best parameter options as obtained from the model training

Optimal parameters ntree == 500, mtry==8  

```{r}
rf_full <- randomForest(return_customer~.,data=train,ntree=500,mtry=8,importance=T, do.trace=100)

res.rf_full <- results(rf_full)

res.rf_full$finalscore
res.rf_full$auc

```


With variables selected from Varimp function

```{r}
rf_selected<- randomForest(frml_rfvars,data=train,ntree=500,mtry=7,importance=T, do.trace=100)

res.rf_selected <- results(rf_selected)

res.rf_selected$finalscore
res.rf_selected$auc
```

Saving results

```{r}
save(rf.caret,file = "rfcaretres.RDS")
```

